Different rules for binocular combination of luminance in cortical and subcortical pathways

### 2022-09-11 manuscript.Rmd
Update manuscript.Rmd

---
title: "manuscript"
author: "Federico Segala, Aurelio Bruno, Alex Wade, Daniel Baker"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Abstract

## Introduction

## Methods

### Participants

### Apparatus & Stimuli

Probably wants a figure here showing example stimuli, maybe some sine waves, and perhaps some stills of large and small pupils taken through the eyetracker (I think we used these on the ECVP poster last year).

### Procedure

### Data analysis

## Results

### Experiment 1

I'd suggest the following figures:

Summary of the pupillometry data with Fourier spectra, maybe an average waveform, and contrast response functions

Summary of the first harmonic EEG data showing the same as for the pupils, maybe with embedded scalp plots inset in the Fourier spectrum

Summary of the second harmonic EEG data showing the same as for the pupils, again with embedded scalp plots inset in the Fourier spectrum

```{r}

# example of how to embed R code in the markdown document

a <- 1

@@

### Experiment 2

>>>>

---
title: "Different rules for binocular combination of luminance in cortical and subcortical pathways"
author: "Federico G. Segala, Aurelio Bruno, Alex R. Wade &  Daniel H. Baker"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    fig_caption: yes
    toc: no
    keep_tex: yes
  word_document: default
  pdf_document:
    toc: no
  html_document: default
bibliography: references.bib
csl: elife.csl
---

```{r setup, include=FALSE}

processdata <- 0  # this flag determines the amount of processing, with 4 levels:
# 0 - do no processing, generate the pdf using existing versions of all figures
# 1 - generate figures using the processed group data (requires XGB of storage)
# 2 - average data across participants, do bootstrapping and modelling
# 3 - download all raw data and analyse (requires XGB of storage)

# reasonably compact code to check which packages are installed, install the missing ones, and activate all
packagelist <- c('knitr','remotes','tictoc','R.matlab','bookdown','grImport','tiff','pals','ez','gtools','signal','boot') # list of CRAN packages
missingpackages <- packagelist[!packagelist %in% installed.packages()[,1]]
if (length(missingpackages)>0){install.packages(missingpackages)}
if (!'osfr' %in% installed.packages()[,1]){remotes::install_github("centerforopenscience/osfr")}
if (!'FourierStats' %in% installed.packages()[,1]){remotes::install_github("bakerdh/FourierStats")}
packagelist <- c(packagelist,'osfr','FourierStats')
toinstall <- packagelist[which(!packagelist %in% (.packages()))]
invisible(lapply(toinstall,library,character.only=TRUE))

knitr::opts_chunk$set(echo = TRUE)

# helper function to make colours transparent
addalpha <- function(col, alpha=1){apply(sapply(col, col2rgb)/255, 2, function(x) rgb(x[1], x[2], x[3], alpha=alpha))}

v4Interp <- function(df, xo, yo, rmax = .75, gridRes = 67) {
  ## Create a function to perform Matlab's v4 interpolation.
  ## Takes as input a data-frame with columns x, y, and z (x co-ordinates, y co-ordinates, and amplitude)
  ## and variables xo and yo, the co-ordinates which will be use to create a grid for interpolation
  xo <- matrix(rep(xo,length(yo)),nrow = length(xo),ncol = length(yo))
  yo <- t(matrix(rep(yo,length(xo)),nrow = length(yo),ncol = length(xo)))
  xy <- df$x + df$y*sqrt(as.complex(-1))
  d <- matrix(rep(xy,length(xy)),nrow = length(xy), ncol = length(xy))
  d <- abs(d - t(d))
  diag(d) <- 1
  g <- (d^2) * (log(d)-1)   # Green's function.
  diag(g) <- 0
  weights <- qr.solve(g,df$z)
  xy <- t(xy)
  outmat <- matrix(nrow = gridRes,ncol = gridRes)
  for (i in 1:gridRes){
    for (j in 1:gridRes) {
      test4 <- abs((xo[i,j] + sqrt(as.complex(-1))*yo[i,j]) - xy)
      g <- (test4^2) * (log(test4)-1)
      outmat[i,j] <- g %*% weights}}
  outDf <- data.frame(x = xo[,1],outmat)
  names(outDf)[1:length(yo[1,])+1] <- yo[1,]
  return(outDf)}

colpal <- c('#FE5000','#8783CF','#228B22','#808080','#6d008b','#8B8000')
targetelectrodes <- c('Oz','POz','O1','O2')

localdir <- 'local/'    # all files are stored in the project directory /local/ which git is told to ignore
if (!file.exists(localdir)){dir.create(localdir)}   # create a local directory to store data and outputs
rawdir <- 'local/rawdata/'
if (!file.exists(rawdir)){dir.create(rawdir)}   # create a local directory to store raw data
matchdir <- 'local/Matching/'
if (!file.exists(matchdir)){dir.create(matchdir)}   # create a local directory to store matching data
TFdir <- 'local/TFdata/'
if (!file.exists(TFdir)){dir.create(TFdir)}   # create a local directory to store  data from the temporal frequency experiment
figdir <- 'local/Figures/'
if (!file.exists(figdir)){dir.create(figdir)}   # create a local directory to store figures
datadir <- 'local/Data/'
if (!file.exists(datadir)){dir.create(datadir)}   # create a local directory to store processed data

# cut off for data inclusion (Mahalanobis distance for complex data)
SDthresh <- 3
nbootstraps <- 1000

@@

# Abstract

# Introduction

# Methods

## Participants

## Apparatus & Stimuli

## Procedure

## Data analysis

@Baker2021

# Results

## Experiment 1

```{r include=FALSE, results='hide'}

# (download and) analyse individual participant data for experiment 1

if (processdata > 2){

legaltriggers <- 1

timeseq <- seq(1/120,10,length.out=120*10)
timeseq2 <- seq(1/120,14,length.out=120*14)
EEGtimes <- seq(1/1000,10,1/1000)
targetindex <- (2*10)+1
maskindex <- (1.6*10)+1
targetindex2 <- (2*2*10)+1
maskindex2 <- (2*1.6*10)+1

  for (participant in 1:30){

    # only download and process data if this has not been done already for this participant
    if (!file.exists(paste0(datadir,'P',100+participant,'summary.RData'))){

      # if there is no directory for this participant's data
      if (!file.exists(paste0(rawdir,'P',100+participant))){
        dir.create(paste0(rawdir,'P',100+participant))

       d <- dir(paste0(rawdir,'P',100+participant))

       if (length(d)==0){

         # if the tar file of this participant's data doesn't exist
         if (!file.exists(paste0(rawdir,'P',100+participant,'.tar'))){
           # download it from OSF

         }

         # then unzip the tar file
         untar(paste0(rawdir,'P',100+participant,'.tar'),exdir=paste0(rawdir,'P',100+participant))

         # and delete the tar file to save storage space
         file.remove(paste0(rawdir,'P',100+participant,'.tar'))
       }

      d <- dir(paste0(rawdir,'P',100+participant),full.names=TRUE)

      pupiltargets <- array(0,dim=c(3,2,60))
pupilmasks <- pupiltargets
pupilwaveforms <- array(0,dim=c(3,2,60,120*14))
pupilspectra <- array(0,dim=c(3,2,60,300))
EEGtargets <- array(0,dim=c(3,64,60))
EEGmasks <- EEGtargets
EEGtargets2 <- EEGtargets
EEGmasks2 <- EEGtargets
EEGwaveforms <- array(0,dim=c(3,64,60,1000*14))
EEGspectra <- array(0,dim=c(3,64,60,300))

psychopyfiles <- dir(path=paste0(rawdir,'P',100+participant),pattern=paste0('P',100+participant,'_CRFstudy'), full.names = TRUE)
infofiles <- dir(path=paste0(rawdir,'P',100+participant),pattern='*_info.csv', full.names = TRUE)
EEGfiles <- dir(path=paste0(rawdir,'P',100+participant),pattern='*_EEG.csv.gz', full.names = TRUE)
pupilfiles <- dir(path=paste0(rawdir,'P',100+participant),pattern='*_pupil_positions.csv', full.names = TRUE)

for (block in 1:length(EEGfiles)){

    infofile <- read.csv(infofiles[block])
  pupilstarttime <- as.numeric(as.character(infofile[4,2]))
  localstarttime <- as.numeric(as.character(infofile[5,2]))

  psychoutput <- read.csv(psychopyfiles[block])

  trialtimes <- psychoutput$trialonset - pupilstarttime
  condorder <- psychoutput$condition

    EEGdata <- read.csv(EEGfiles[block])
  electrodes <- colnames(EEGdata)

      triggertimes <- NULL
    counter <- 0
    lasttrigger <- -10000
    for (n in 1:nrow(EEGdata)){
      if(EEGdata$Trigger[n] %in% legaltriggers){
        if(n>(lasttrigger+10000)){
        counter <- counter + 1
        triggertimes[counter] <- n
        lasttrigger <- n
      }}
    }

    for (cond in 1:60){
      for (ch in 1:64){
        trial <- EEGdata[(triggertimes[cond]+(2001:12000)),ch+2]
        fspec <- (fft(trial)/length(trial))
        EEGtargets[block,ch,condorder[cond]] <- fspec[targetindex]
        EEGmasks[block,ch,condorder[cond]] <- fspec[maskindex]
        EEGtargets2[block,ch,condorder[cond]] <- fspec[targetindex2]
        EEGmasks2[block,ch,condorder[cond]] <- fspec[maskindex2]
        EEGspectra[block,ch,condorder[cond],1:300] <- fspec[1:300]

        trial <- EEGdata[(triggertimes[cond]+(-999:13000)),ch+2]
        EEGwaveforms[block,ch,condorder[cond],] <- trial - mean(trial[1:1000])
      }}

  pdata <- read.csv(pupilfiles[block])
  pdata2 <- pdata[,c(1,3,4,14)]
  pdata2[,1] <- pdata2[,1] - localstarttime
  pdata2 <- pdata2[which(pdata2[,3]>0.0),]

  for (eye in 1:2){
    eyedata <- pdata2[which(pdata2[,2]==(2-eye)),]
    for (cond in 1:60){
      a <- which(eyedata[,1]>trialtimes[cond]+2)
      b <- which(eyedata[,1]<(trialtimes[cond]+12))
      i <- intersect(a,b)
      trial <- eyedata[i,]
      trial[,1] <- trial[,1] - trial[1,1]
      if (nrow(trial)>3){
        resampled <- interp1(trial[,1],trial[,4],timeseq,method='linear',extrap=TRUE)
        fspec <- (fft(resampled)/length(resampled))
        pupiltargets[block,eye,condorder[cond]] <- fspec[targetindex]
        pupilmasks[block,eye,condorder[cond]] <- fspec[maskindex]
        pupilspectra[block,eye,condorder[cond],1:300] <- fspec[1:300]
      }

      a <- which(eyedata[,1]>(trialtimes[cond]-1))
      b <- which(eyedata[,1]<(trialtimes[cond]+13))
      i <- intersect(a,b)
      trial <- eyedata[i,]
      trial[,1] <- trial[,1] - trial[1,1]
      if (nrow(trial)>3){
        resampled <- interp1(trial[,1],trial[,4],timeseq2,method='linear',extrap=TRUE)
        pupilwaveforms[block,eye,condorder[cond],] <- resampled - mean(resampled[1:120])
      }

      }}

}

cleanmeansP <- matrix(0,nrow=6,ncol=5)
cleanmasksP <- matrix(0,nrow=6,ncol=5)
cleanmeansE <- array(0,dim=c(64,6,5))
cleanmasksE <- array(0,dim=c(64,6,5))
cleanmeansE2 <- array(0,dim=c(64,6,5))
cleanmasksE2 <- array(0,dim=c(64,6,5))

for (cond in 1:6){
  startindex <- (10*(cond-1))
  for (level in 1:5){
    temp <- c(pupiltargets[,,startindex+level],pupiltargets[,,startindex+level+5])
        cartdata <- data.frame(Re(temp),Im(temp))
    if (sum(cartdata)!=0){
      D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
      i <- which(D<SDthresh) 
    cleanmeansP[cond,level] <- mean(temp[i])}

    temp <- c(pupilmasks[,,startindex+level],pupiltargets[,,startindex+level+5])
        cartdata <- data.frame(Re(temp),Im(temp))
    if (sum(cartdata)!=0){
      D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
      i <- which(D<SDthresh) 
    cleanmasksP[cond,level] <- mean(temp[i])}

    for (ch in 1:64){
    temp <- c(EEGtargets[,ch,startindex+level],EEGtargets[,ch,startindex+level+5])
            cartdata <- data.frame(Re(temp),Im(temp))
    if (sum(cartdata)!=0){
      D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
      i <- which(D<SDthresh) 
    cleanmeansE[ch,cond,level] <- mean(temp[i])}

    temp <- c(EEGmasks[,ch,startindex+level],EEGmasks[,ch,startindex+level+5])
                cartdata <- data.frame(Re(temp),Im(temp))
    if (sum(cartdata)!=0){
      D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
      i <- which(D<SDthresh) 
    cleanmasksE[ch,cond,level] <- mean(temp[i])}

    temp <- c(EEGtargets2[,ch,startindex+level],EEGtargets2[,ch,startindex+level+5])
                cartdata <- data.frame(Re(temp),Im(temp))
    if (sum(cartdata)!=0){
      D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
      i <- which(D<SDthresh) 
    cleanmeansE2[ch,cond,level] <- mean(temp[i])}

    temp <- c(EEGmasks2[,ch,startindex+level],EEGmasks2[,ch,startindex+level+5])
                cartdata <- data.frame(Re(temp),Im(temp))
    if (sum(cartdata)!=0){
      D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
      i <- which(D<SDthresh) 
    cleanmasksE2[ch,cond,level] <- mean(temp[i])}   
    }

  }
}

electrodeindices <- match(targetelectrodes,electrodes)-2

meanwavesP <- apply(pupilwaveforms[,,c(15,20),],4,mean,na.rm=TRUE)
meanwavesE <- apply(EEGwaveforms[,electrodeindices,c(15,20),],4,mean,na.rm=TRUE)
meanspectraP <- apply(pupilspectra[,,c(15,20),],4,mean,na.rm=TRUE)
meanspectraE <- apply(EEGspectra[,electrodeindices,c(15,20),],4,mean,na.rm=TRUE)
save(file=paste(datadir,'P',100+participant,'summary.RData',sep=''),list=c('cleanmeansP','cleanmeansE','cleanmeansE2','cleanmasksP','cleanmasksE','cleanmasksE2','meanspectraP','meanspectraE','meanwavesP','meanwavesE','electrodes'))

      }

      }

    }
  }

```{r include=FALSE, results='hide'}

# load in individual participant data and average, save group data

if (processdata > 1){

hdata <- read.csv(paste0(EEGdir,'headerfile.csv'))

d <- dir(datadir,pattern='*summaryC.RData',full.names=TRUE)

tempmasksP <- array(0,dim=c(length(d),6,5))
tempmeansP <- tempmasksP
tempmasksE <- array(0,dim=c(length(d),64,6,5))
tempmeansE <- tempmasksE
tempmeansE2 <- tempmeansE
tempmasksE2 <- tempmasksE
tempspectraP <- array(0,dim=c(length(d),300))
tempspectraE <- array(0,dim=c(length(d),300))
tempwavesP <- array(0,dim=c(length(d),1680))
tempwavesE <- array(0,dim=c(length(d),14000))

  for (s in 1:length(d)){

    load(d[s])

  tempmasksP[s,,] <- cleanmasksP
  tempmeansP[s,,] <- cleanmeansP
  tempspectraP[s,] <- meanspectraP
  tempspectraE[s,] <- meanspectraE
  tempwavesP[s,] <- meanwavesP
  tempwavesE[s,] <- meanwavesE
  tempmeansE[s,,,] <- cleanmeansE
  tempmasksE[s,,,] <- cleanmasksE
  tempmeansE2[s,,,] <- cleanmeansE2
  tempmasksE2[s,,,] <- cleanmasksE2

  }

electrodeindices <- match(targetelectrodes,electrodes)-2

tempspectraP <- tempspectraP[-c(2,4,13,20,26),]

cleanmeansP <- matrix(0,nrow=6,ncol=5)
cleanmeansPCI <- array(0,c(2,6,5))
cleanmasksP <- matrix(0,nrow=6,ncol=5)
cleanmasksPCI <- array(0,c(2,6,5))
cleanmeansEav <- matrix(0,nrow=6,ncol=5)
cleanmeansEavCI <- array(0,c(2,6,5))
cleanmasksEav <- matrix(0,nrow=6,ncol=5)
cleanmasksEavCI <- array(0,c(2,6,5))
cleanmeansE2av <- matrix(0,nrow=6,ncol=5)
cleanmeansE2avCI <- array(0,c(2,6,5))
cleanmasksE2av <- matrix(0,nrow=6,ncol=5)
cleanmasksE2avCI <- array(0,c(2,6,5))
cleanmeansE <- array(0,dim=c(64,6,5))
cleanmeansECI <- array(0,c(2,64,6,5))
cleanmasksE <- array(0,dim=c(64,6,5))
cleanmasksECI <- array(0,c(2,64,6,5))
cleanmeansE2 <- array(0,dim=c(64,6,5))
cleanmeansE2CI <- array(0,c(2,64,6,5))
cleanmasksE2 <- array(0,dim=c(64,6,5))
cleanmasksE2CI <- array(0,c(2,64,6,5))

for (cond in 1:6){
  for (level in 1:5){

    temp <- tempmeansP[,cond,level]
    cartdata <- data.frame(Re(temp),Im(temp))
    if (sum(cartdata)!=0){
      D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
      i <- which(D<SDthresh) 
    mbs <- amperrors(temp[i],method='boot',quantiles=95,nresamples=nbootstraps)
    cleanmeansP[cond,level] <- mbs$meanamp
    cleanmeansPCI[1,cond,level] <- mbs$lowerCI
    cleanmeansPCI[2,cond,level] <- mbs$upperCI
    }

    temp <- tempmasksP[,cond,level]
    cartdata <- data.frame(Re(temp),Im(temp))
    if (sum(cartdata)!=0){
      D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
      i <- which(D<SDthresh)

          mbs <- amperrors(temp[i],method='boot',quantiles=95,nresamples=nbootstraps)
    cleanmasksP[cond,level] <- mbs$meanamp
    cleanmasksPCI[1,cond,level] <- mbs$lowerCI
    cleanmasksPCI[2,cond,level] <- mbs$upperCI
}

      temp <- apply(tempmeansE[,electrodeindices,cond,level],1,mean)
      cartdata <- data.frame(Re(temp),Im(temp))
      if (sum(cartdata)!=0){
        D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
        i <- which(D<SDthresh) 
    mbs <- amperrors(temp[i],method='boot',quantiles=95,nresamples=nbootstraps)
    cleanmeansEav[cond,level] <- mbs$meanamp
    cleanmeansEavCI[1,cond,level] <- mbs$lowerCI
    cleanmeansEavCI[2,cond,level] <- mbs$upperCI
      }

temp <- apply(tempmasksE[,electrodeindices,cond,level],1,mean)
      cartdata <- data.frame(Re(temp),Im(temp))
      if (sum(cartdata)!=0){
        D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
        i <- which(D<SDthresh) 
mbs <- amperrors(temp[i],method='boot',quantiles=95,nresamples=nbootstraps)
    cleanmasksEav[cond,level] <- mbs$meanamp
    cleanmasksEavCI[1,cond,level] <- mbs$lowerCI
    cleanmasksEavCI[2,cond,level] <- mbs$upperCI
        }

temp <- apply(tempmeansE2[,electrodeindices,cond,level],1,mean)
      cartdata <- data.frame(Re(temp),Im(temp))
      if (sum(cartdata)!=0){
        D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
        i <- which(D<SDthresh) 
    mbs <- amperrors(temp[i],method='boot',quantiles=95,nresamples=nbootstraps)
    cleanmeansE2av[cond,level] <- mbs$meanamp
    cleanmeansE2avCI[1,cond,level] <- mbs$lowerCI
    cleanmeansE2avCI[2,cond,level] <- mbs$upperCI
        }

      temp <- apply(tempmasksE2[,electrodeindices,cond,level],1,mean)
      cartdata <- data.frame(Re(temp),Im(temp))
      if (sum(cartdata)!=0){
        D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
        i <- which(D<SDthresh) 
    mbs <- amperrors(temp[i],method='boot',quantiles=95,nresamples=nbootstraps)
    cleanmasksE2av[cond,level] <- mbs$meanamp
    cleanmasksE2avCI[1,cond,level] <- mbs$lowerCI
    cleanmasksE2avCI[2,cond,level] <- mbs$upperCI
        }

    for (ch in 1:64){

      temp <- tempmeansE[,ch,cond,level]
      cartdata <- data.frame(Re(temp),Im(temp))
      if (sum(cartdata)!=0){
        D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
        i <- which(D<SDthresh) 
    mbs <- amperrors(temp[i],method='boot',quantiles=95,nresamples=nbootstraps)
    cleanmeansE[ch,cond,level] <- mbs$meanamp
    cleanmeansECI[1,ch,cond,level] <- mbs$lowerCI
    cleanmeansECI[2,ch,cond,level] <- mbs$upperCI
}

      temp <- tempmasksE[,ch,cond,level]
      cartdata <- data.frame(Re(temp),Im(temp))
      if (sum(cartdata)!=0){
        D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
        i <- which(D<SDthresh) 
mbs <- amperrors(temp[i],method='boot',quantiles=95,nresamples=nbootstraps)
    cleanmasksE[ch,cond,level] <- mbs$meanamp
    cleanmasksECI[1,ch,cond,level] <- mbs$lowerCI
    cleanmasksECI[2,ch,cond,level] <- mbs$upperCI
        }

      temp <- tempmeansE2[,ch,cond,level]
      cartdata <- data.frame(Re(temp),Im(temp))
      if (sum(cartdata)!=0){
        D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
        i <- which(D<SDthresh) 
    mbs <- amperrors(temp[i],method='boot',quantiles=95,nresamples=nbootstraps)
    cleanmeansE2[ch,cond,level] <- mbs$meanamp
    cleanmeansE2CI[1,ch,cond,level] <- mbs$lowerCI
    cleanmeansE2CI[2,ch,cond,level] <- mbs$upperCI
        }

      temp <- tempmasksE2[,ch,cond,level]
      cartdata <- data.frame(Re(temp),Im(temp))
      if (sum(cartdata)!=0){
        D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
        i <- which(D<SDthresh) 
    mbs <- amperrors(temp[i],method='boot',quantiles=95,nresamples=nbootstraps)
    cleanmasksE2[ch,cond,level] <- mbs$meanamp
    cleanmasksE2CI[1,ch,cond,level] <- mbs$lowerCI
    cleanmasksE2CI[2,ch,cond,level] <- mbs$upperCI
        }

    }}}

meanspectraP <- NULL
meanspectraE <- NULL
meanspectraPCI <- array(0,c(2,300))
meanspectraECI <- array(0,c(2,300))

for (f in 1:300){

mbs <- amperrors(tempspectraP[,f],method='boot',quantiles=95,nresamples=nbootstraps)
meanspectraP[f] <- mbs$meanamp
meanspectraPCI[1,f] <- mbs$lowerCI
meanspectraPCI[2,f] <- mbs$upperCI

mbs <- amperrors(tempspectraE[,f],method='boot',quantiles=95,nresamples=nbootstraps)
meanspectraE[f] <- mbs$meanamp
meanspectraECI[1,f] <- mbs$lowerCI
meanspectraECI[2,f] <- mbs$upperCI

}

meanwavesP <- apply(tempwavesP,2,mean)
meanwavesE <- apply(tempwavesE,2,mean)
meanwavesPCI <- array(0,dim=c(2,length(meanwavesP)))
meanwavesECI <- array(0,dim=c(2,length(meanwavesE)))
nsubjs <- dim(tempwavesP)[1]
for (t in 1:length(meanwavesP)){
  bspop <- NULL
  for (s in 1:nbootstraps){bspop[s] <- mean(sample(tempwavesP[,t],nsubjs,replace=TRUE))}
  meanwavesPCI[,t] <- quantile(bspop,c(0.025,0.975))
}
for (t in 1:length(meanwavesE)){
  bspop <- NULL
  for (s in 1:nbootstraps){bspop[s] <- mean(sample(tempwavesE[,t],nsubjs,replace=TRUE))}
  meanwavesECI[,t] <- quantile(bspop,c(0.025,0.975))
}

save(file=paste0(datadir,'Averagedata.RData'),list=c('cleanmeansP','cleanmeansPCI','cleanmasksP','cleanmasksPCI','cleanmeansEav','cleanmeansEavCI','cleanmasksEav','cleanmasksEavCI','cleanmeansE2av','cleanmeansE2avCI','cleanmasksE2av','cleanmasksE2avCI','cleanmeansE','cleanmeansECI','cleanmasksECI','cleanmeansE2','cleanmeansE2CI','cleanmasksE2','cleanmasksE2CI','meanspectraP','meanspectraPCI','meanspectraE','meanspectraECI','meanwavesP','meanwavesE','meanwavesPCI','meanwavesECI','electrodes'))

}

```{r include=FALSE, results='hide'}

# generate plot showing the group pupillometry data

if (processdata > 0){

contrastsdB <- 20*log10(c(6,12,24,48,96))

load(file=paste0(datadir,'Averagedata.RData'))

plotlims <- c(12,40,0,0.03)  # define the x and y limits of the plot (minx,maxx,miny,maxy)
ticklocsx <- contrastsdB    # locations of tick marks on x axis
ticklocsy <- seq(0,0.03,0.01)    # locations of tick marks on y axis
ticklabelsx <-c(6,12,24,48,96)        # set labels for x ticks
ticklabelsy <- ticklocsy    # set labels for y ticks

postscript("CRF1p.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 5.5)

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   # create an empty axis of the correct dimensions
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1) 
title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)    # titles for axes
title(ylab="Amplitude at 2Hz (mm)", col.lab=rgb(0,0,0), line=1.8, cex.lab=1.5)

for (cond in 1:3){
  lines(contrastsdB,abs(cleanmeansP[cond,]), col=colpal[cond], lwd=3, cex=0.5)
  arrows(contrastsdB,abs(cleanmeansP[cond,]),contrastsdB,cleanmeansPCI[1,cond,],lwd=3,col=colpal[cond],length=0.05,angle=90)
  arrows(contrastsdB,abs(cleanmeansP[cond,]),contrastsdB,cleanmeansPCI[2,cond,],lwd=3,col=colpal[cond],length=0.05,angle=90)
  points(contrastsdB,abs(cleanmeansP[cond,]), pch = 20+cond, col='black', bg=colpal[cond], cex=1.6, lwd=3)   # draw the data points themselves
}  
legend(19,0.03,c('Monocular','Binocular','Dichoptic'),pch=21:23,pt.bg=colpal[1:3],pt.lwd=3,pt.cex=1.6,box.lwd=2)
text(14,0.028,'(c)',adj=0.5,cex=2.5)

dev.off()

postscript("CRF2p.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 5.5)

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   # create an empty axis of the correct dimensions
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)   
title(ylab="Amplitude at 2Hz (mm)", col.lab=rgb(0,0,0), line=1.8, cex.lab=1.5)

condlist <- c(1,5,6)
pointlist <- c(21,24,25)
for (cond in 1:3){
  lines(contrastsdB,abs(cleanmeansP[condlist[cond],]), col=colpal[condlist[cond]], lwd=3, cex=0.5)  
  arrows(contrastsdB,abs(cleanmeansP[condlist[cond],]),contrastsdB,cleanmeansPCI[1,condlist[cond],],lwd=3,col=colpal[condlist[cond]],length=0.05,angle=90)
  arrows(contrastsdB,abs(cleanmeansP[condlist[cond],]),contrastsdB,cleanmeansPCI[2,condlist[cond],],lwd=3,col=colpal[condlist[cond]],length=0.05,angle=90)  
  points(contrastsdB,abs(cleanmeansP[condlist[cond],]), pch = pointlist[cond], col='black', bg=colpal[condlist[cond]], cex=1.6, lwd=3)   
}  
legend(19,0.03,c('Monocular','Binocular cross','Dichoptic cross'),pch=pointlist,pt.bg=colpal[condlist],pt.lwd=3,pt.cex=1.6,box.lwd=2)
text(14,0.028,'(d)',adj=0.5,cex=2.5)

dev.off()

postscript("CRF3p.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 5.5)

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   # create an empty axis of the correct dimensions
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)    # titles for axes
title(ylab="Amplitude at 1.6Hz (mm)", col.lab=rgb(0,0,0), line=1.8, cex.lab=1.5)

for (cond in 4:6){
  lines(contrastsdB,abs(cleanmasksP[cond,]), col=colpal[cond], lwd=3, cex=0.5)   
  arrows(contrastsdB,abs(cleanmasksP[cond,]),contrastsdB,cleanmasksPCI[1,cond,],lwd=3,col=colpal[cond],length=0.05,angle=90)
  arrows(contrastsdB,abs(cleanmasksP[cond,]),contrastsdB,cleanmasksPCI[2,cond,],lwd=3,col=colpal[cond],length=0.05,angle=90)    
  points(contrastsdB,abs(cleanmasksP[cond,]), pch = pointlist[cond-3], col='black', bg=colpal[cond], cex=1.6, lwd=3) 
}
legend(19,0.03,c('Monocular 1.6Hz','Binocular cross','Dichoptic cross'),pch=pointlist,pt.bg=colpal[4:6],pt.lwd=3,pt.cex=1.6,box.lwd=2)
text(14,0.028,'(e)',adj=0.5,cex=2.5)

dev.off()

postscript("SpecP.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 8)

frequencies <- (0:299)/10

plotlims <- c(1,6,0,0.03) 
ticklocsx <- seq(1,6,1)    # locations of tick marks on x axis
ticklocsy <- seq(0,0.03,0.01)    # locations of tick marks on y axis
ticklabelsx <-ticklocsx        # set labels for x ticks
ticklabelsy <- ticklocsy    # set labels for y ticks

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   # create an empty axis of the correct dimensions
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)  
title(xlab="Frequency (Hz)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)  
title(ylab="Amplitude (mm)", col.lab=rgb(0,0,0), line=1.8, cex.lab=1.5)

polygon(frequencies[c(11:61,61:11)],c(meanspectraPCI[1,11:61],meanspectraPCI[2,61:11]),col=colpal[2],border=NA)

lines(frequencies[11:61],abs(meanspectraP[11:61]), col=colpal[2], lwd=3, cex=0.5)

text(1.1,0.028,'(b)',adj=0.5,cex=2.5)

dev.off()

postscript("timecourseP.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 8)

times <- seq((1/120)-1,13,1/120)
sinewave <- sin(2*times * 2*pi)
sinewave[1:120] <- 0
sinewave[1560:1680] <- 0

plotlims <- c(-1,13,-1,1) 
ticklocsx <- seq(-1,13,1)    # locations of tick marks on x axis
ticklocsy <- seq(-1,1,1)    # locations of tick marks on y axis
ticklabelsx <- ticklocsx        # set labels for x ticks
ticklabelsy <- ticklocsy    # set labels for y ticks

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)  
title(xlab="Time (s)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)  
title(ylab="Amplitude (mm)", col.lab=rgb(0,0,0), line=1.8, cex.lab=1.5)
